---
title: "Masters Homework 2"
author: "Fred LaPolla"
date: "6/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
##test

## On Your Own Work

>- When you read in a .csv file into R, what class is the object that you read in (like nyc)?

#### it is a data.frame

>- What command can you use to determine what type of data is in a column?

```{r}
sapply(nyc, class) #shows the type for each column in a dataframe 

class(nyc$CHOLRAT) #shows only the class of the specified column if you know the name
```


>- Provide a command to find the type of data in the DX_DBTS column of nyc. 

```{r}
class(nyc$DX_DBTS)
```


>- Provide the same command to find the type of data but use the other way of indexing from what you did above. 

```{r}
sapply(nyc[10], class)
```



>- What would happen if we had chosen the first 4 columns of NYC HANES and made a matrix? What would the class of the cells in this matrix be? Why?


#### In a matrix all data.types must be the same, so converting the first four columns into a matrix would make the class of all the cells 'characters' and any column that is different would become "N.A"



>- If we make a vector,vec<- c(3,9,4,7,8,9), and multiply it by WannaMatrix, what is the value in the third row, third column? 

```{r}
WannaMatrix <- nyc[5:10, 6:9]
WannaMatrix <- as.matrix(WannaMatrix,nrow = 6, ncol = 4)

vec<- c(3,9,4,7,8,9)

value <- vec*WannaMatrix
print(value[3,3])

```



>- What happens if instead we made a vector vec2 <- c(3,9,4,7,8) and multiply it by WannaMatrix, what is the third row, third column value? What is going on? 

```{r}
vec2 <- c(3,9,4,7,8)

value2 <- vec2*WannaMatrix
```
#### This multiplication doesn't work because vec2 has only 5 variables but WannaMatrix has 6 rows. Mutliplying vec with the Matrix multiplies each row with the corresponding integer in the vector, since the new vector isn't a multiple of 6 it cannot be used with the matrix. 

>- Look up the arguments for read.csv

```{r}
?read.csv
```


>- Use apply() to find the median value of CADMIUM, LEAD CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, and CREATININE

```{r}
apply(nyc[,12:19],2, median, na.rm = TRUE)
```


>- Do the same using sapply()

```{r}
sapply(nyc[,12:19], median, na.rm = TRUE)
```


>- Use a for loop to find the median value of columns 16:19

```{r}
for (i in 16:19){
  med_val <- median(nyc[,i], na.rm = TRUE)
  print(c(colnames(nyc[i]), med_val))
}
```


>- Make a for loop to find the standard deviation (hint sd() ) of CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, AND CREATININE

```{r}
for (i in 16:19){
  sd_val <- sd(nyc[,i], na.rm = TRUE)
  print(c(colnames(nyc[i]), sd_val))
}
```


>- Bonus/Challenge Create a vector in which to put the standard deviations 

```{r}
sd_vec <- c()

for (i in 16:19){
  sd_val <- sd(nyc[,i], na.rm = TRUE)
  sd_vec <- append(sd_vec, sd_val)
}
print(sd_vec)

#In the "tips" from the lecture you said not to grow objects in a loop-- does this include append?
```


>- Use ifelse to create a new variable of high and normal A1C dichotomizing with 7 as a threshold for high levels

```{r}

nyc$highA1C <- ifelse(nyc$A1C>=7, 0, 1)
nyc$highA1C <- factor(nyc$highA1C)
summary(nyc$highA1C)
```

